---
title: "Analysis of Opinions on Women's Role in Caregiving and Labour with 2021 GSS Data"
author: "Eshta Bhardwaj"
thanks: "Code and data available at: https://github.com/eshtab/GSSdata2021_analysis"
date: "April 21 2023"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
number-sections: true
format: 
  pdf:
    fig-pos: 'H' #this  makes sure that figures are placed directly underneath code
bibliography: "references.bib"
toc: true
---


```{r}
#| include: FALSE
#| warning: FALSE
#| message: FALSE

#### Workspace set-up ####

library(tidyverse)
library(janitor)
library(modelsummary)
library(lubridate)

```

# Introduction

The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.

# Data Collection 

## Survey Design

## Non-response and Imputation

# Data Description

## Overview

## Data Cleaning

## Variables of Interest

```{r}
#| include: FALSE
#| warning: FALSE
#| message: FALSE


# read in data from inputs folder of R project
cleaned_data <-
  read_csv(
    file = "./../../inputs/data/cleaned_data.csv",
    show_col_types = FALSE,
  )
```

```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE

# summary visualizations for each variable

# sex
pie(table(cleaned_data$sex), 
    labels = paste(round(prop.table(table(cleaned_data$sex))*100), "%", sep = ""), 
    col = rainbow(2),
    main = "Sex at Birth")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE

# age
cleaned_data |>
  ggplot(mapping = aes(x = age)) +
  geom_bar() +
  labs(
    x = "Age",
    y = "Number of Responses"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +

  scale_fill_brewer(palette = "Set1")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# marital status
cleaned_data |>
  ggplot(mapping = aes(x = marital_status)) +
  geom_bar() +
  labs(
    x = "Marital Status",
    y = "Number of Responses"
  ) +
  theme_classic() +
  scale_fill_brewer(palette = "Set1")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# make degree a  factored variable to enable ordering
cleaned_data <-
  cleaned_data |>
  mutate(
    degree = factor(degree),
    degree = fct_relevel(
      degree,
      "Less than High School",
      "High School",
      "Associate/Junior College",
      "Bachelors",
      "Graduate",
      NA
    )
  )
# degree
cleaned_data |>
  ggplot(mapping = aes(x = degree)) +
  geom_bar() +
  labs(
    x = "Degree",
    y = "Number of Responses"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +

  scale_fill_brewer(palette = "Set1")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# occ_prestige
hist(cleaned_data$occ_prestige,
     xlab = 'Prestige Score', ylab = 'Number of Responses', main = 'Occupational Prestige Scores')

# skipping industry
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# num_children
cleaned_data |>
  ggplot(mapping = aes(x = num_children)) +
  geom_bar() +
  labs(
    x = "Number of Children",
    y = "Number of Responses"
  ) +
  theme_classic() +
  scale_fill_brewer(palette = "Set1")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# op working mother
pie(table(cleaned_data$op_working_mother), 
    labels = paste(cleaned_data$op_working_mother, sep =  " ", 
                   round(prop.table(table(cleaned_data$op_working_mother))*100), "%"), 
                   main = "Opinion on whether 'A working mother can establish just as warm and secure a
relationship with her children as a mother who does not work'.",
    col = rainbow(n_distinct(cleaned_data$op_working_mother)))
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# op preschool
pie(table(cleaned_data$op_preschool), 
    labels = paste(cleaned_data$op_preschool, sep =  " ", 
                   round(prop.table(table(cleaned_data$op_preschool))*100), "%"), 
                   main = "Opinion on whether 'A preschool child is likely to suffer if his or her mother works'.",
    col = rainbow(n_distinct(cleaned_data$op_preschool)))
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# op family
pie(table(cleaned_data$op_family), 
    labels = paste(cleaned_data$op_family, sep =  " ", 
                   round(prop.table(table(cleaned_data$op_family))*100), "%"), 
                   main = "Opinion on whether 'It is much better for everyone involved if the man is the
achiever outside the home and the woman takes care of the
home and family'.",
    col = rainbow(n_distinct(cleaned_data$op_family)))
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# op meovrwrk
pie(table(cleaned_data$op_men_over_work), 
    labels = paste(cleaned_data$op_men_over_work, sep =  " ", 
                   round(prop.table(table(cleaned_data$op_men_over_work))*100), "%"), 
                   main = "Opinion on whether 'Family life often suffers because men concentrate too much on their work'.",
    col = rainbow(n_distinct(cleaned_data$op_men_over_work)))
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# op childcare
pie(table(cleaned_data$op_childcare), 
    labels = paste(cleaned_data$op_childcare, sep =  " ", 
                   round(prop.table(table(cleaned_data$op_childcare))*100), "%"), 
                   main = "Opinion on whether 'We are spending too much, too little, or about the right amount on) Assistance for childcare'.",
    col = rainbow(n_distinct(cleaned_data$op_childcare)))
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# work status
cleaned_data |>
  ggplot(mapping = aes(x = work_status)) +
  geom_bar() +
  labs(
    x = "Work Status",
    y = "Number of Responses"
  ) +
  scale_x_discrete(labels = function(x) 
    stringr::str_wrap(x, width = 6)) +
  #theme(axis.text.x = element_text(angle = 20, hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
# hours_worked
cleaned_data |>
  ggplot(mapping = aes(x = hours_worked)) +
  geom_bar() +
  labs(
    x = "Number of Hours Worked",
    y = "Number of Responses"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  scale_fill_brewer(palette = "Set1")
```


```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE

# make ever_worked a  factored variable to enable ordering
cleaned_data <-
  cleaned_data |>
  mutate(
    ever_worked = factor(ever_worked),
    ever_worked = fct_relevel(
      ever_worked,
      "YES",
      "NO",
      "Inapplicable",
      "No answer",
      "Skipped on Web",
    )
  )
# ever_worked
cleaned_data |>
  ggplot(mapping = aes(x = ever_worked)) +
  geom_bar() +
  labs(
    x = "Answers to 'Ever Worked?'",
    y = "Number of Responses"
  ) +
  theme_classic() +
  scale_fill_brewer(palette = "Set1")

```

# Results

```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE

# Look at all opinions broken down by men vs women 

# make op_working_mother a factored variable to enable ordering
cleaned_data <-
  cleaned_data |>
  mutate(
    op_working_mother = factor(op_working_mother),
    op_working_mother = fct_relevel(
      op_working_mother,
      "Strongly Agree",
      "Agree",
      "Disagree",
      "Strongly Disagree",
      NA
    )
  )

# op_working_mother
cleaned_data |>
  ggplot(mapping = aes(x = op_working_mother)) +
  geom_bar() +
  labs(
    x = "Opinion on whether 'A working mother can establish just as warm and secure a
relationship with her children as a mother who does not work'",
    y = "Number of Responses"
  ) +
  theme_classic() +
  # https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_brewer(palette = "Set1") +
  # subplots by sex at birth
  facet_wrap(
    ~sex,
    scales = "free_x"
  )

```


# Discussion
- note some opinions have 4 scale others have 5

# Conclusion

\newpage

\appendix

# Appendix  {.unnumbered}

# Survey
\newpage

# References